>

### 2024-11-25 docs/exec-summary/_preprocessing.qmd
Draft the initial results
Explain the weighting method used for adjusting DDD #23

Daily medication claims is a tabular data containing values from single medication entity for each patients in the cohort. Data transformation was performed to construct a square matrix of aggregate medication claim per day, with rows and columns representing groups of medication. Each element in the upper and lower triangle represents co-prescriptions, while the diagonal of the matrix represents single prescriptions. All elements in the matrix is a tally of drug combination/single prescription in a population level claimed on a daily basis. The matrix follows a knowledge graph construct with a subject-predicate-object (s, p, o) triples, i.e. `Medication1-PrescribedAlongside-Medication2`. The knowledge graph is a representation of drug prescription network (DPN), where the edge (predicate) is the number of concurrency. In the original DPN, the edge weight was the number of claims by assuming that all of defined daily doses (DDD) were equal to 1 [@Cavallo2012]. To account for augmentation polypharmacy, DDD provides further information due to the presence of a sub-dose medication, where DDD < 1 should be allowed. Naturally, an augmentation polypharmacy regimen should have a higher weight since this regimen is intended to increase the efficacy of the augmented medication [@Shrivastava2013]. Therefore, weight was re-applied on the edges by dividing the number of claims by the DDD.

$$
\stackrel{\textrm{Unregularized}}{
  \begin{bmatrix}
  n_{1, 1} & \dots & n_{1, N} \\
  \vdots & \ddots & \vdots \\
  n_{N, 1} & \dots & n_{N, N} \\
  \end{bmatrix}
}
\to
\stackrel{\textrm{Regularized by } DDD}{
  \begin{bmatrix}
  \frac{1}{2} \cdot \left(\frac{n_1}{DDD_1} + \frac{n_1}{DDD_1} \right) & \dots & \frac{1}{2} \cdot \left(\frac{n_1}{DDD_1} + \frac{n_N}{DDD_N} \right) \\
  \vdots & \ddots & \vdots \\
  \frac{1}{2} \cdot \left(\frac{n_N}{DDD_N} + \frac{n_1}{DDD_1} \right) & \dots & \frac{1}{2} \cdot \left(\frac{n_N}{DDD_N} + \frac{n_N}{DDD_N} \right) \\
  \end{bmatrix}
}
$$ {#eq-regularize-mtx}

>>>>

Daily medication claims is a tabular data containing values from single medication entity for each patients in the cohort. Data transformation was performed to construct a square matrix of aggregate medication claim per day, with rows and columns representing groups of medication. Each element in the upper and lower triangle represents co-prescriptions, while the diagonal of the matrix represents single prescriptions. All elements in the matrix is a tally of drug combination/single prescription in the population claimed on a daily basis. The matrix follows a graph construct, where we can denote the row-to-column as `Medication1-PrescribedAlongside-Medication2`. The graph is a representation of drug prescription network (DPN), where the edge is the number of concurrency. In the original DPN, the edge weight was the number of claims by assuming that all of defined daily doses (DDD) were equal to 1 [@Cavallo2012]. However, not all medications were prescribed with DDD equals to 1, and discarding these entries would lead to loss of information. Therefore, we applied weights based on DDD, where the weight is the highest if DDD equals to 1, and gradually diminishes as it is getting further from 1. To adjust the weight, first we set the baseline weight as $\omega_B = 1$. Then, we calculated the weight based on the expected density of a normal distribution, where the expected value is $\mu = \omega_B$ and standard deviation is $\sigma = \frac{\omega_B}{3}$. The expected value was set to $\omega_B$ because we expect entries with DDD equals to 1. The standard deviation was set to $\frac{\omega_B}{3}$ to represent an equidistant deviation from the expected value to 0. As a result, we obtained the weighted DDD as $\omega$.

$$\omega = \frac{1}{\frac{\omega_B}{3} \sqrt{2 \pi}} e^{-\frac{1}{2}\left(\frac{DDD - \omega_B}{\omega_B/3}\right)^2}$$ {#eq-weighted-DDD}

$$
\stackrel{\textrm{Unregularized}}{
  \begin{bmatrix}
  n_{1, 1} & \dots & n_{1, N} \\
  \vdots & \ddots & \vdots \\
  n_{N, 1} & \dots & n_{N, N} \\
  \end{bmatrix}
}
\to
\stackrel{\textrm{Regularized by } DDD}{
  \begin{bmatrix}
  \frac{1}{2} \cdot \left(\omega_1 + \omega_1 \right) & \dots & \frac{1}{2} \cdot \left(\omega_1 + \omega_N \right) \\
  \vdots & \ddots & \vdots \\
  \frac{1}{2} \cdot \left(\omega_N + \omega_1 \right) & \dots & \frac{1}{2} \cdot \left(\omega_N + \omega_N \right) \\
  \end{bmatrix}
}
$$ {#eq-regularize-mtx}
